```{r fig-5-setup}
knitr::opts_chunk$set(
  fig.path = "./fig-5/"
)
```

## Main Figures

### THIQ vs aMSH meta-regression

```{r amsh-vs-thiq-meta, fig.width=6, fig.height=6, cache=TRUE}
amsh_vs_thiq_wide <- cre %>%
  filter(dose == "low") %>%
  select(pos, aa, compound, est = log2FoldChange, se = std.error) %>%
  pivot_wider(names_from = compound, values_from = c(est, se)) %>%
  mutate(
    contrast = est_aMSH - est_THIQ,
    pooled_se = sqrt(se_aMSH^2 + se_THIQ^2),
    z = contrast / pooled_se,
    pval = (1 - pnorm(abs(z))) * 2,
    fdr = p.adjust(pval, method = "fdr"),
    mutant = str_c(pos, aa)
  )

# assume a perfect slope and intercept centered at 0
priors <- c(
  prior(normal(1, 1), class = "b", lb = 0),
  prior(normal(0, 1), class = "Intercept")
)

bench::bench_time(
  model_brm <- brm(
    data = amsh_vs_thiq_wide,
    formula = est_aMSH | se(se_aMSH) ~ me(est_THIQ, se_THIQ),
    backend = "cmdstanr",
    chains = 4, cores = 4,
    prior = priors,
    family = "gaussian"
  )
)

meta_lm <- fixef(model_brm)[, 1]

resid_data <- residuals(model_brm) %>%
  as_tibble() %>%
  mutate(
    residual_z = Estimate / Est.Error,
    residual_pval = (1 - pnorm(abs(residual_z))) * 2,
    residual_fdr = p.adjust(residual_pval, method = "fdr")
  ) %>%
  bind_cols(amsh_vs_thiq_wide)

resid_data %>%
  ggplot(aes(y = est_aMSH, x = est_THIQ)) +
  geom_hline(yintercept = 0, lty = "dashed") +
  geom_vline(xintercept = 0, lty = "dashed") +
  geom_abline(lty = "dashed") +
  geom_point(pch = 21, alpha = 0.5) +
  geom_point(
    color = "#1F77B4",
    data = filter(resid_data, residual_fdr < 0.05, residual_z < 0),
    size = 2
  ) +
  geom_point(
    color = "#FF7F0E",
    data = filter(resid_data, residual_fdr < 0.05, residual_z > 0),
    size = 2
  ) +
  geom_abline(
    intercept = meta_lm[1],
    slope = meta_lm[2],
    color = "#E377C2"
  ) +
  geom_text_repel(
    aes(label = mutant),
    data = filter(resid_data, residual_fdr < 0.05, (residual_z > 0)),
    force = 2,
    min.segment.length = 0,
    max.overlaps = 20,
    max.time = 2
  ) +
  coord_equal() +
  labs(
    x = "THIQ\n(log2-fold change from WT)",
    y = "aMSH\n(log2-fold change from WT)",
    title = "aMSH vs THIQ (low concentration)"
  ) +
  theme(panel.grid.major = element_blank())
```

As a table, here's the 5% FDR's

```{r}
resid_data %>%
  filter(residual_fdr < 0.05) %>%
  select(pos, aa, est_aMSH, est_THIQ, residual_z, residual_fdr) %>%
  mutate(residual_fdr = -log10(residual_fdr)) %>%
  arrange(pos, residual_z) %>%
  knitr::kable(digits = 2)
```

### aMSH vs THIQ -- Position 129 

```{r pos129-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq <- function(sel_pos) {
  aa_ordering <- tribble(
    ~aa, ~group,
    "*", "Special",
    "C", "Special",
    "G", "Special",
    "P", "Special",
    "A", "Hydrophobic",
    "V", "Hydrophobic",
    "I", "Hydrophobic",
    "L", "Hydrophobic",
    "M", "Hydrophobic",
    "F", "Hydrophobic",
    "Y", "Hydrophobic",
    "W", "Hydrophobic",
    "R", "Positive",
    "H", "Positive",
    "K", "Positive",
    "D", "Negative",
    "E", "Negative",
    "S", "Polar Uncharged",
    "T", "Polar Uncharged",
    "N", "Polar Uncharged",
    "Q", "Polar Uncharged"
  ) %>%
    mutate(group = factor(group, levels = c("Special", "Hydrophobic", "Polar Uncharged", "Positive", "Negative")))

  cre %>%
    filter(
      pos == sel_pos,
      dose == "low"
    ) %>%
    mutate(aa = as.character(aa)) %>%
    inner_join(aa_ordering) %>%
    mutate(aa = factor(aa, levels = aa_ordering$aa)) %>%
    ggplot(aes(x = aa, y = log2FoldChange, color = compound, group = compound)) +
    geom_hline(yintercept = 0, lty = "dashed") +
    geom_pointrange(
      aes(
        ymin = log2FoldChange - 2 * std.error,
        ymax = log2FoldChange + 2 * std.error
      ),
      position = position_dodge(width = 0.5)
    ) +
    facet_grid(cols = vars(group), space = "free_x", scales = "free_x") +
    labs(
      title = str_c("Pos ", sel_pos, " Mutants"),
      y = "Log2-fold Change vs WT\n(+/- 2 Standard Errors)",
    ) +
    scale_color_manual(values = c("#1F77B4", "#FF7F0E")) +
    theme(
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.grid.major.y = element_line(colour = "grey92", linewidth = rel(0.5))
    )
}

compare_amsh_thiq(129) +
  scale_y_continuous(limits = c(-3, 1.25), breaks = seq(-3, 1, 1))
```

### aMSH vs THIQ -- Position 284

```{r pos284-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq(284) +
  scale_y_continuous(limits = c(-1.75, 0.75), breaks = seq(-1.5, 0.5, 0.5), expand = c(0, 0))
```

## Extended Figures

### PCA for ligand selectivity

```{r ligand-pca, fig.width=9, fig.height=9}
ligand_pca <- cre %>%
  filter(compound != "None") %>%
  mutate(condition = str_c(compound, dose, sep = " - ")) %>%
  select(pos, aa, condition, statistic) %>%
  pivot_wider(names_from = condition, values_from = statistic) %>%
  unite(col = "variant", pos, aa, sep = "") %>%
  column_to_rownames("variant") %>%
  prcomp()

ligand_pca %>%
  fviz_pca(
    axes = c(1, 3),
    label = "var",
    ggtheme = theme_get()
  ) +
  geom_point(
    aes(x = .fittedPC1, y = .fittedPC3),
    size = 2,
    data = augment(ligand_pca) %>% filter(.fittedPC3 < -3.75),
    color = "#FF7F0E"
  ) +
  geom_point(
    aes(x = .fittedPC1, y = .fittedPC3),
    size = 2,
    data = augment(ligand_pca) %>% filter(.fittedPC3 > 3.75),
    color = "#1F77B4"
  ) +
  geom_text_repel(
    aes(x = .fittedPC1, y = .fittedPC3, label = .rownames),
    data = augment(ligand_pca) %>% filter(abs(.fittedPC3) > 3.75),
    min.segment.length = 0,
    max.time = 2
  ) +
  labs(
    title = "PCA Biplot of aMSH and THIQ conditions"
  )
```

### aMSH vs THIQ -- Position 48

```{r pos48-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq(48) +
  scale_y_continuous(limits = c(-1.75, 0.75), breaks = seq(-1.5, 0.5, 0.5), expand = c(0, 0))
```

### DRC aMSH vs Thiq -- Position 48

```{r 48-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all <- function(sel_pos) {
  amsh_zero <- cre %>%
    filter(
      pos == sel_pos,
      compound %in% c("None", "aMSH")
    ) %>%
    mutate(compound = "aMSH")

  thiq_zero <- cre %>%
    filter(
      pos == sel_pos,
      compound %in% c("None", "THIQ")
    ) %>%
    mutate(compound = "THIQ")

  bind_rows(
    amsh_zero,
    thiq_zero,
  ) %>%
    ggplot(aes(x = dose, y = log2FoldChange, color = compound, group = compound)) +
    geom_hline(yintercept = 0, lty = "dashed") +
    geom_pointrange(
      aes(
        ymin = log2FoldChange - 2 * std.error,
        ymax = log2FoldChange + 2 * std.error
      ),
      position = position_dodge(width = 0.25)
    ) +
    geom_line(
      position = position_dodge(width = 0.25)
    ) +
    facet_wrap(facets = vars(aa), nrow = 5) +
    theme(
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.grid.major.y = element_line(colour = "grey92", linewidth = rel(0.5))
    ) +
    scale_color_manual(values = c("#1F77B4", "#FF7F0E")) +
    labs(
      title = paste0(sel_pos, " Dose Response Curves"),
      y = "Log2-fold Change vs WT",
      x = "Concentration",
      color = "Drug"
    )
}

compare_amsh_thiq_all(48) +
  scale_y_continuous(limits = c(-2, 1), breaks = seq(-2, 1, 0.5), expand = c(0, 0))
```

### aMSH vs THIQ -- Position 52

```{r pos52-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq(52)
```

### DRC aMSH vs THIQ -- Position 52

```{r 52-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all(52) +
  scale_y_continuous(limits = c(-2, 1), breaks = seq(-2, 1, 0.5), expand = c(0, 0))
```

### aMSH vs THIQ -- Position 101

```{r pos101-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq(101)
```

### DRC aMSH vs THIQ -- Position 101

```{r 101-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all(101)
```

### aMSH vs THIQ -- Position 104

```{r pos104-amsh-vs-thiq, fig.width=9, fig.height=4}
compare_amsh_thiq(104)
```

### DRC aMSH vs THIQ -- Position 104

```{r 104-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all(104)
```

### DRC aMSH vs THIQ -- Position 129

```{r 129-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all(129)
```

### DRC aMSH vs THIQ -- Position 284

```{r 284-drc, fig.width=12, fig.height=12}
compare_amsh_thiq_all(284)
```
